---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# symbiontmodeler

<!-- badges: start -->
<!-- badges: end -->

The goal of `symbiontmodeler` is to implement various models for symbiont frequency dynamics. Applications to [Wolbachia](https://www.nature.com/articles/nrmicro1969) frequency dynamics are emphasized, but the models may also be applied to other CI-causing symbionts like Cardinium.  

## Installation

You can install the development version of `symbiontmodeler` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jmgraham30/symbiontmodeler")
```

## Background

Following

Turelli, M., Katznelson, A., & Ginsberg, P. S. (2022). Why Wolbachia-induced cytoplasmic incompatibility is so common.*Proceedings of the National Academy of Sciences*, 119(47), e2211637119.

and

Hoffmann, A. A., Turelli, M., & Harshman, L. G. (1990). Factors affecting the distribution of cytoplasmic incompatibility in Drosophila simulans. *Genetics*, 126(4), 933-948.

we implement the frequency dynamics defined by

$$p_{t+1} = \frac{p_{t}F(1 - \mu)}{1 + p_{t}(F - 1 - s_{h}) + p_{t}^{2}s_{h}(1 - \mu F)},$$
where

* $p_{t}$ is the frequency of symbiont infected females in the population at time $t$,

* $F$ is the relative fecundity of symbiont infected females,

* $\mu$ is the transmission rate, and

* $s_{h}$ is the cytoplasmic incompatibility (CI) rate.

For a finite population of size $N$, the iteration is given by

$$
\begin{align}
p_{t+1} &= \frac{p_{t}F(1 - \mu)}{1 + p_{t}(F - 1 - s_{h}) + p_{t}^{2}s_{h}(1 - \mu F)},  \\
q_{ij} &= \left(\begin{array}{c} N \\ j \end{array}\right)(p^{\ast})^{j}(1 - p^{\ast})^{N - j}. 
\end{align}
$$



## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(symbiontmodeler)
## basic example code
```

```{r examp_sims, warning=FALSE, message=FALSE}
library(ggplot2)
library(patchwork)

tp_1 <- turelli_simulation(0.4,1.2,0.1,0.0,1000)
p_1 <- plot_a_simulation(tp_1,alpha_val=0.2) + ylim(c(0,1))

tp_2 <- turelli_simulation(0.4,1.02,0.5,0.0,1000)
p_2 <- plot_a_simulation(tp_2,alpha_val=1.0) + ylim(c(0,1))

tp_3 <- turelli_simulation(0.1,1.02,0.1,0.8,1000)
p_3 <- plot_a_simulation(tp_3,alpha_val=0.8) + ylim(c(0,1))

tp_4 <- turelli_simulation(0.3,1.02,0.1,0.8,1000)
p_4 <- plot_a_simulation(tp_4,alpha_val=0.2) + ylim(c(0,1))

tp_5 <- infection_freq_rmu_ci_iteration(1.225,c(0.05,0.8),c(0.99,0.01),0.1,10000)
p_5 <- plot_a_simulation(tp_5,alpha_val=0.2) + ylim(c(0,1))

tp_6 <- infection_freq_rf_rmu_iteration(1.2,0.1,c(0.01,0.0),c(1.0,0.0),1000)
p_6 <- plot_a_simulation(tp_6,alpha_val=0.2) + ylim(c(0,1))

(p_1 + p_2) / (p_3 + p_4) / (p_5 + p_6)
```


